The socialization of meritocracy and market justice preferences at school

### 2024-08-06 07-supplementary-material.qmd
Archivos submit
Push formato
Co-Authored-By: Kevin Carrasco <67479225+Kevin-carrasco@users.noreply.github.com>

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
---

# Supplementary Material

@@

# Translated items for market justice preferences

@@

# Complete multilevel models

@@

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
#| label: tbl-lineal-reg
ccoef <- list(
 "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  universitarioSi = "University or posgraduate",
  "universitarioNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  acc_tec = "Technology acces",
  internet_rec2 = "No internet connectivity (Ref.= Internet connectivity)",
  school_nseMedio = "SES Medium",
  school_nseAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High",
  mean_educ = "Prop. university level at school",
  "cod_depe2Part. subvencionado" = "Subsidized Private",
  "cod_depe2Part. privado" = "Private")

>>>>

# Supplementary material

@@

## Translated items for market justice preferences

@@

## Complete multilevel models

@@

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
#| label: tbl-lineal-reg-complete
ccoef <- list(
 "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  universitarioSi = "University or posgraduate",
  "universitarioNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  acc_tec = "Technology acces",
  internet_rec2 = "No internet connectivity (Ref.= Internet connectivity)",
  school_nseMedio = "SES Medium",
  school_nseAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High",
  mean_educ = "Prop. university level at school",
  "cod_depe2Part. subvencionado" = "Subsidized Private",
  "cod_depe2Part. privado" = "Private")

### 2024-08-06 paper-blinded.qmd
Push formato
Co-Authored-By: Kevin Carrasco <67479225+Kevin-carrasco@users.noreply.github.com>

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
abstract: |
  Previous research has shown that schools justify student performance differences based on meritocratic ideals. However, the link between school meritocracy and students’ justification of inequalities has received little attention. This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile. Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools. Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults. Most students endorse meritocratic views, particularly the reward of effort. Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity. At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement. Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance. These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
keywords: 
  - market justice 
  - meritocracy
  - socialization
  - family
  - schools
  - Chile
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

@@

>>>>

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  mdpi-pdf:
    keep-tex: true
mdpi_journal: behavsci
article_type: article
article_status: submit
#author:
#  - name: Juan Carlos Castillo
#    corresponding: true
#    orcid: 0000-0003-1265-7854
#    email: juancastillov@uchile.cl
#    role:
#      - Conceptualization
#      - Supervision
#      - Methodology
#      - Writing - review and editing
#    affiliations:
#      - name: "Universidad de Chile"
#        department: Department of Sociology
#        id: 1
#        country: Chile
#  - name: Mauricio Salgado
#    orcid: 0000-0002-8108-6154
#    role:
#      - Conceptualization
#      - Methodology
#      - Writing - review and editing    
#    affiliations:
#      - name: Centro de Estudios Públicos
#        id: 2
#        department: CEP
#        country: Chile
#  - name: Kevin Carrasco
#    orcid: 0000-0002-9348-8361
#    role:
#      - Methodology
#      - formal analysis
#      - software
#      - visualization
#      - Writing - review and editing
#    affiliations:
#      - name: Center for Social Conflict and Cohesion Studies
#        id: 3
#        department: COES
#        country: Chile
#  - name: Andreas Laffert
#    orcid: 0000-0002-9008-2454
#    role:
#      - Methodology
#      - formal analysis
#      - software
#      - visualization
#      - Writing - review and editing  
#    affiliations:
#      - name: Universidad Católica de Chile
#        id: 4
#        department: Institute of Sociology
#        country: Chile
fundingstatement: | 
  The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
dataavailability: |
  The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>
  Access to the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>
conflictsofinterest: |
  The authors declare no conflict of interest.
abstract: |
  Previous research has shown that schools justify student performance differences based on meritocratic ideals. However, the link between school meritocracy and students’ justification of inequalities has received little attention. This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile. Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools. Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults. Most students endorse meritocratic views, particularly the reward of effort. Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity. At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement. Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance. These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
keywords: [market justice, meritocracy, socialization, family, schools, Chile]
bibliography: "input/bib/just-ineq-merit.bib"
---

@@

{{< include 07-supplementary-material.qmd >}}
